%!TEX root = ./jctt.tex

\section{Conclusions and Future Work}
We have presented the VEF method for one-group neutron transport in slab geometry and the pairing of Lumped Linear Discontinuous Galerkin for the \SN transport step and the constant-linear Mixed Finite Element Method for the drift diffusion acceleration step. We have numerically demonstrated that the LLDG/MFEM VEF method accelerates Source Iteration by transferring the rapid convergence of the angular shape of the angular flux to the scalar flux. In addition, the VEF method performed similarly to consistently differenced S$_2$SA. 

Methods for increased consistency between LLDG and MFEM were also presented. This included a more consistent method for computing the Eddington factor on the cell centers and a cell centered slope reconstruction. It was shown that both the VEF method without and without added consistency measureswere second-order accurate as expected from the orders of accuracy of LLDG and MFEM in isolation and that all of the VEF methods were robust in the diffusion limit. In addition, while this nonlinear scheme produces two solutions, one from \SN and one from drift diffusion, the solutions were shown to converge as the mesh was refined for both homogeneous and inhomogeneous systems. 

Slope reconstruction significantly increased the convergence between the \SN and drift diffusion solutions. While consistent representation of the Eddington factor did not significantly impact numerical accuracy or solution convergence, it is expected to be important in radiative transfer calculations due to the non-linearities in temperature and the spatial dependence of the cross sections. 

The VEF method is especially suited for multiphysics calculations as VEF drift diffusion is conservative and much less expensive than a transport sweep. In addition, the multiphysics and transport discretizations can be arbitrarily different. 

% add rational polynomial conclusions: expected to be important in rad transfer 
% mfem is conservative and can be paired to multiphysics 
% take out consistency talk 

Future work includes extending the VEF method presented in this paper to the radiative transfer equations, verifying the VEF method in higher dimensions, and generalizing the method to higher order finite elements. 